---
title: "Final Project - SOLUTIONS"
output: 
  html_document:
    df_print: paged
    theme: journal
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set( echo = FALSE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

```{r, eval=TRUE, message=FALSE, warning=FALSE}

# clear the workspace
rm( list = ls() )

# load the libraries you will need
library( network )
library( sna )

# Load the networks

# set the url
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-PHX-arrest-2023-net.rds"

# for the 2023
PhxArrestNet2023 <- readRDS( url( loc ) )

# set the url
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-PHX-arrest-2022-net.rds"

# for the 2022
PhxArrestNet2022 <- readRDS( url( loc ) )

```

```{r setup-functions, echo = FALSE, eval = TRUE}

# This section sets up the functions we will use in the report

# node rescale function
rescale <- function( nchar, low, high ){
  min_d <- min( nchar )
  max_d <- max( nchar )
  rscl  <- ( ( high - low )*( nchar - min_d ) ) / ( max_d - min_d ) + low
  rscl
}

# edge rescale function
edge.rescale <- function( uniMat, low, high ){
  diag( uniMat ) <- 0
  min_w <- min( uniMat[uniMat != 0] )
  max_w <- max( uniMat[uniMat != 0] )
  rscl <- ( ( high-low )  * ( uniMat[uniMat != 0] - min_w ) ) / ( max_w - min_w ) + low
  rscl
}

# edge shading fuction
edge.shade <- function( uniMat ){
  net.edges <- edge.rescale( uniMat, 0.01, 1 )
  vec.to.color <- as.vector( abs( net.edges ) )
  vec.to.color <- 1 - vec.to.color # subtract 1 to flip the grey function scale.
  edge.cols <- grey( vec.to.color )
  return( edge.cols )
}


```

## Part I: Two-Mode Network

```{r part-I-object-setup, echo=FALSE, eval=TRUE}

# This builds all the objects I want to use in the report.
# They are built here, but the code does not show in the report.

# ---- 
# calculate the density of the network

PhxArrestNet2023Mat <- as.matrix( PhxArrestNet2023 ) # create a matrix from the network object
L <- sum( PhxArrestNet2023Mat )                      # identify the number of edges in the graph
N <- dim( PhxArrestNet2023Mat )[1]                   # identify the number of actors in the example
M <- dim( PhxArrestNet2023Mat )[2]                   # identify the number of events in the example
density.PhxArrestNet2023Mat <- L / ( N * M )         # calculate the density


# ----
# calculate the degree centrality scores for each set of nodes  

actor.deg <- rowSums( PhxArrestNet2023Mat ) # raw scores for actors
event.deg <- colSums( PhxArrestNet2023Mat ) # raw scores for events

# calculate the mean degree centrality scores
mean.actor.deg <- L / N # mean degree for actors
mean.event.deg <- L / M # mean degree for events


# ----
# calculate the standardized degree centrality scores for each set of nodes  
# we need these for the comparisons in part III

s.actor.deg <- actor.deg / M # standardized score for actors
s.event.deg <- event.deg / N # standardized score for events

```

### *What can you tell me about this network?* 

```{r part-I-net, echo=FALSE, eval=TRUE, fig.width=10, fig.height=10, fig.fullwidth=TRUE}

# set the plot regions to ease with visualization
par( 
  mfrow = c( 1, 1 ),
  mar = c( 2, 1, 4, 1)
  )

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PhxArrestNet2023,                        # our network to plot
  gmode = "twomode",                       # indicate it is two modes
  usearrows = FALSE,                       # turn off the arrowheads
  main="Phoenix Co-Arrest Network (2023)", # add a title
  vertex.col = cbind( 
    rep( "#34ebcf", dim(PhxArrestNet2023Mat )[1] ),
    rep( "#484a69", dim(PhxArrestNet2023Mat )[2] ) ),
  vertex.cex = rescale( cbind( actor.deg, event.deg ), 0.2, 3 ) # size the nodes
)

```

#### Density

The density of for the `PhxArrestNet2023` network indicates that `r round( density.PhxArrestNet2023Mat, 3 ) * 100` percent of the ties that could be observed, were observed. This also means that `r 1 - round( density.PhxArrestNet2023Mat, 3 ) * 100` of the ties that could have occurred were not observed. Overall, we can say that the network is fairly *sparse*.

<br>

#### Degree Centrality

For the first set of nodes, the mean degree is `r round( mean.actor.deg, 2 )`, meaning that on average, individuals were arrested `r round( mean.actor.deg, 2 )` times together. For the second set of nodes, the incidents where individuals were arrested together, the mean degree is `r round( mean.event.deg, 2 )`, meaning that each incident, on average, had `r round( mean.event.deg, 2 )` individuals arrested. The variation in these scores is visualized in the plot below:  

```{r part-I-bar-plots, echo=FALSE, eval=TRUE, fig.width=10, fig.height=5, fig.fullwidth=TRUE}

# set the plot regions to ease with visualization
par( 
  mfrow = c( 1, 2 ),
  mar = c( 2, 1, 4, 1)
  )

# plot the degree distributions
hist( actor.deg,
      main = "Degree Distribution for Arrestees",
      col = "#34ebcf",
      breaks = 20
      )

hist( event.deg,
      main = "Degree Distribution for Incidents",
      breaks = 20,
      col = "#484a69"
      )

```

<br>

----

<br>

## Part II: One-Mode Network

```{r part-II-object-setup, echo=FALSE, eval=TRUE}

# This builds all the objects I want to use in Part II of the report.
# They are built here, but the code does not show in the report.

# ---- 
# create the projections

# create the "person" matrix
PhxArrestNet2023MatP <- PhxArrestNet2023Mat %*% t( PhxArrestNet2023Mat )


# ----
# create the raw centrality scores

# degree
deg.arrest <- degree( PhxArrestNet2023MatP, gmode = "graph", cmode = "degree" )

# closeness
g <- dim( as.matrix( PhxArrestNet2023MatP ) )[1]
close.arrest <- closeness( PhxArrestNet2023MatP, gmode="graph" ) / ( g - 1 )

# betweenness
btwn.arrest <- betweenness( PhxArrestNet2023MatP, gmode="graph" )


# ----
# create the standardized centrality scores

s.deg.arrest <- deg.arrest / ( g - 1 ) # standardized score is deg / g-1 where g is the number of nodes
s.close.arrest <- closeness( PhxArrestNet2023MatP, gmode="graph" ) # standardized closeness is the score reported by default
s.btwn.arrest <- btwn.arrest / ( ( ( g-1 ) * ( g-2 ) ) / 2 ) # standardized betweenness is the betweenness / ( ( ( g-1 ) * ( g-2 ) ) / 2 )


# ----
# mean degree and standardized degree

mean.deg.arrest    <- mean( deg.arrest )
mean.close.arrest  <- mean( close.arrest )
mean.btwn.arrest   <- mean( btwn.arrest )
mean.s.deg.arrest   <- mean( s.deg.arrest )
mean.s.close.arrest <- mean( s.close.arrest )
mean.s.btwn.arrest  <- mean( s.btwn.arrest )


# ----
# calculate the graph centralization for degree, closeness, and betweenness centrality

dcent.arrest <- centralization( PhxArrestNet2023MatP, degree, mode="graph", cmode="degree" )
ccent.arrest <- centralization( PhxArrestNet2023MatP, closeness, mode="graph" )
bcent.arrest <- centralization( PhxArrestNet2023MatP, betweenness, mode="graph" )


```

### *Who are the central actors in this network?* 

```{r part-II-net, echo=FALSE, eval=TRUE, fig.width=10, fig.height=10, fig.fullwidth=TRUE}

# render the plot
gplot( 
  PhxArrestNet2023MatP,
  gmode = "graph",
  vertex.col = "#34ebcf",
  main = "One Mode Projection of Co-Arrests (Person Matrix) Network"
  )



```

#### Centrality

In summarizing the centrality scores for the network, we can say the following:  

  * The mean *degree* centrality score is `r mean.deg.arrest` meaning that the average node has `r mean.deg.arrest` edges incident on them.   
  * The mean *closeness* centrality score is `r mean.close.arrest`, meaning that the average node is very close to all other nodes in the network.   
  * Finally, the mean betweenness centrality score is `r mean.btwn.arrest`, meaning that the average node sits on `r mean.btwn.arrest` paths between *i* and *j*.   

To visualize these differences, the plot below shows the co-arrest network where the nodes are sized by degree, closeness, and betweenness centrality scores, respectively.

```{r part-II-net-cent, echo=FALSE, eval=TRUE, fig.width=10, fig.height=10, fig.fullwidth=TRUE}

# create the inputs for the loop
titles  <- c( "Degree", "Closeness", "Betweenness")
cols    <- c( "#5c1233", "#0c7091", "#949c7b" )
v.sizes <- cbind( deg.arrest, close.arrest, btwn.arrest )

# change the margins using the mar= argument
# change the number of plots using the mfrow= argument
par( 
  mfrow=c( 2,2 ),
  mar=c( 0.1, 0.5, 1, 0.1 ) 
  )

for( i in 1:3 ){
  
  # set the seed
  set.seed( 507 )

  # render the plot
  gplot( 
  PhxArrestNet2023MatP,
  gmode = "graph",
  vertex.col = cols[i],
  vertex.cex = rescale( v.sizes[,i], 0.5, 1.5 ),
  main = titles[i]
  )

}

```

#### Centralization

Regarding the centralization of the network for each centrality score:  

  * The *degree* centralization score is `r round( dcent.arrest, 2 )` meaning that...  
  * The *closeness* centralization score is `r round( ccent.arrest, 2 )` ...
  * The *betweenness* centralization score is `r round( bcent.arrest, 2 )` ... 

As these summaries indicate...COMPARISON

```{r, echo = FALSE}

# reset the plot layout
par( mfrow=c( 2,2 ) )

# clear the workspace
rm( list = ls() )

```

<br>

----

<br>

## Part III: Comparison

*How has the co-offending network changed from 2022 to 2023?* 

YOUR ANSWER TO THIS QUESTION HERE (THEN DELETE THIS LINE)

```{r, echo=FALSE}

# your code here for creating objects here

```
